robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack

### 2023-07-14 paper_elsevier/paper_elsevier.Rmd
update robs

NOAA ObsPack data has been used to support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases Observing 
SATellite (GOSAT) data. 
@saunois2016global 
At regional scale, @lu2022methane performed another studied focused
on north america using as priors local emissions inventories. 
Furthermore, @hu2023trend presented trends and cycles of methane over the 
US.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrfats is different than surface stations, but all the files
include concentrations and other critical fields. 
An important information is the scale of measurement, which isXX.
Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `robspack` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `robspack`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focued on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output and plots.

>>>>

NOAA ObsPack data has been used to support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases 
Observing SATellite (GOSAT) data. 
@saunois2016global.
At regional scale, @lu2022methane performed another studied focused
on north america using as priors local emissions inventories. 
Furthermore, @hu2023trend presented trends and cycles of methane over the 
US.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. 
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `robspack` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `robspack`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process consists in creating a 
summary of the ObsPack files, reading them in an iteration process, 
filtering, and generating another output and plots.
